Parallel Global Search Algorithm with Local Tuning\\ for Solving Mixed-Integer Global Optimization Problems

### 2021-03-21 2021/Lobachevskii_BarkalovLebedev/ljmsamp.tex
after corrections

The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox. Table \ref{tab:1} shows the number of trials it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}. The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods. All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM. The experimental results show the superiority of the PMGAL method both in terms of the number of iterations and the running time.

\begin{table}[ht]
	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
	\label{tab:1}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{PMGAL} \\
		\cline{2-3} \cline{4-5} 
		 & $k$ & $t$ &  $k$ & $t$  \\
		\hline 
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & 	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & 	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & 	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
		\hline
	\end{tabular}
\end{table}

@@

In total, we generated and solved 100 9-dimensional mixed-integer problems of this type. In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
The accuracy of the search was equal to $\epsilon = 10^{-2}$. Computational experiments were carried out on the Lobachevsky supercomputer. The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM.

@@

\begin{table}
	\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
	\label{tab:3}
	\center
	\begin{tabular}{|c|c|c|c|}
		\hline	
	\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\

		\hline		
		2  & 199.0 & 65.4 & 61.0  \\
		5  & 91.2 & 28.0 &  26.1 \\
		17  & 37.2 & 10.8 & 7.5 \\
		65  & 29.3 & 8.7 & 5.4 \\
		\hline
	\end{tabular}
\end{table}

The last series of experiments was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
Table \ref{tab:3} shows the average solution time  $T_{av}$ (in seconds) depending on the involved number of processes $N_{p}$ and threads $N_{t}$ in each process. Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. 
The acceleration provided by the parallel algorithm is not ideal, but quite acceptable.

@@

This article presents the results related to the development of parallel algorithms for global optimization for solving multiextremal problems in which some of the parameters are continuous, and some are discrete. 
These problems can also include non-convex constraints, which are accounted for using a special index scheme.
We propose a parallel algorithm for solving the problems of this class. The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.

The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak. This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.

In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.

>>>>

The first series of experiments was carried out using a sequential version of the algorithm without using local tuning in order to compare it with other similar methods.
The PMGAL algorithm proposed in this work (in its sequential version) was compared with the genetic algorithm for solving mixed-integer optimization problems implemented in the Matlab Global Optimization Toolbox. Table \ref{tab:1} shows the number of trials and time it took for these methods to solve a number of mixed-integer test problems taken from \cite{Deep,Floudas}.  
%Russian
The test problems used contained 2 continuous and 1 discrete parameter, which had from 3 to 6 values, depending on the specific problem.
Also, the problem statement included from 1 to 4 non-convex constraints.  
Note that the presence of complex constraints that form a non-convex feasible region prevents the use of methods based on the principles of local optimization. 
The genetic algorithm from Matlab Global Optimization Toolbox was used with the recommended parameters (for example, the population size was 50, the maximum number of generations was 200, and the crossover fraction was 0.8). The PMGAL algorithm was used with standard parameters providing convergence to the solution ($r_\nu$ from (\ref{R}) was equal to 2.3 for all $\nu = 1,2,3$, the evolvent construction parameter $M=10$).
The same accuracy of search $\epsilon = 10^{-2}$ was used for both methods.

All computational experiments were carried out on a computer with an Intel Core i5-7300 2.5 GHz processor and 8 Gb RAM. The experimental results show the superiority of the PMGAL method both in terms of the number of iterations and the running time.

\begin{table}[ht]
	\caption{Comparison of the effectiveness of the PMGAL and GA methods}
	\label{tab:1}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c|}{ GA } &  \multicolumn{2}{c|}{PMGAL} \\
		\cline{2-3} \cline{4-5} 
		 & Number of trials & Time, sec. &  Number of trials & Time, sec.  \\
		\hline 
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & 	417 &	0.04 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & 	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & 	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & 	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 &  372	 &	0.017 \\
		\hline
	\end{tabular}
\end{table}

@@

In total, we generated and solved 100 mixed-integer problems of this type with 5 discrete and 4 continuous parameters. In order to simulate the computational complexity inherent in applied optimization problems, the calculation of the function in all the experiments was complicated by additional calculations that did not change the form of the function and the location of its minima.
The accuracy of the search was equal to $\epsilon = 10^{-2}$. Computational experiments were carried out on the Lobachevsky supercomputer. The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs (16 cores in total) and 64 Gb RAM.

@@

%\begin{table}[ht]
	%\caption{Average running time $T_{av}$ of the parallel algorithm with local tuning}
	%\label{tab:3}
	%\center
	%\begin{tabular}{|c|c|c|c|}
		%\hline	
	%\diaghead{\theadfont Diag Column} { $N_{p}$}{$N_{t}$} & 1 & 4 & 8\\	
		%\hline		
		%2  & 199.0 & 65.4 & 61.0  \\
		%5  & 91.2 & 28.0 &  26.1 \\
		%17  & 37.2 & 10.8 & 7.5 \\
		%65  & 29.3 & 8.7 & 5.4 \\
		%\hline
	%\end{tabular}
%\end{table}

The last experiment was performed to evaluate the acceleration of the parallel version of the PMGAL algorithm. 
%Russian
We used the parallel algorithm with previously used parameters to solve the same series of 100 problems with 5 discrete and 4 continuous parameters.
Fig.~\ref{fig:3} shows the speedup $S(p)$ depending on the number of processes and threads involved in each process. Here, the acceleration is the ratio of the running time of the sequential algorithm  $t_1$ to the running time of the parallel algorithm $t_p$, i.e. $S(p) = t_1/t_p$, where $p$ is the total number of threads used.
Our experiments involved from 1 to 16 cluster nodes depending on the number of parallel processes. The speedup provided by the parallel algorithm is not ideal, but quite acceptable.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.6\textwidth]{speedup.pdf}
    \caption{Speedup of the parallel algorithm with local tuning}
    \label{fig:3}
\end{figure}

@@

This article presents the results related to the development of parallel algorithms for global optimization for solving multiextremal problems in which some of the parameters are continuous, and some are discrete. 
These problems can also include non-convex constraints, which are accounted for using a special index scheme.

We propose a parallel algorithm for solving the problems of this class. The basic principle of organizing parallel computations in the algorithm is based on the simultaneous carrying out of several trials at different points of the search domain.
The parallel global optimization algorithm uses a local tuning technique based on the assumption that the multiextremity of the problem is weak. This assumption is true, in particular, at the final phase of the search, when one or more trials have already been performed in the vicinity of the global solution, and the only remaining step is to refine the found estimate of the solution.
In its sequential form, the proposed algorithm remains on par with the similar methods implemented in Matlab Global Optimization Toolbox. 
In a parallel form, the proposed algorithm demonstrates acceptable scalability up to dozens of parallel processes.

%Russian
The principal assumption which ensures that the proposed algorithm is applicable is that the Lipschitz condition for the functions of the problem with varying continuous parameters for each fixed set of discrete parameters is fulfilled.
Optimization problems in which the functions are Lipschitzian belong to the class of computationally expensive problems. Any numerical algorithm for solving such a problem requires the values of the function to be calculated at the nodes of a particular (uniform or non-uniform) grid in the entire domain of admissible parameter changes.
Although the assumption of Lipschitzness makes it possible to estimate the optimum on a substantially non-uniform (i.e., more resource-efficient than uniform) grid in the search domain, the scale of the problems being solved here is rather limited in comparison with the problems for a local extremum. In a reasonable time, even using parallel algorithms, it is possible to solve problems with approximately 10 continuous and 10 discrete parameters. 
A solution to problems with hundreds of variables can be based on a nested optimization scheme assuming that the dependence of the objective function on some of the variables is local \cite{Barkalov2020_1}.

An example of a complex applied problem to which the developed algorithms could be applied is the selection of the optimal design parameters of an aircraft hot-air anti-icing system. The mathematical model of the anti-icing system, along with the parameters that change continuously (for example, the diameter of the tube supplying warm air from the engine to the wing; the diameter of the outlets on this tube; the distance between the outlets; etc.), also contains discrete parameters (for example, the number of rows of outlets, which due to the technical limitations may be equal to 1, 2 or 3). An example of such a problem is featured in \cite{Habashi}. The use of the algorithm proposed in the paper makes it possible to avoid optimization for each value of a discrete parameter separately, and instead cut off unpromising options in advance during the search process.
